Finding Quality Issues in SKOS Vocabularies

### 2012-03-26 analysis.tex
added authoritative concepts comment
added some more analysis data
added some keywords for results interpretation

\todo{BH}{I think we can remove the collections column, none of the issues involve collections}

@@

We used Sindice and available listings\footnote{\url{http://www.w3.org/2001/sw/wiki/SKOS/Datasets}} to learn about existing SKOS vocabularies. Table~\ref{tab:vocabs} lists our representative vocabulary selection and shows that the vocabularies differ in size and application domain: the \emph{Gemeenschappelijke Thesaurus Audiovisuele Archieven (GTAA)} is vocabulary from the media domain, \emph{Eurovoc} and the \emph{Integrated Public Sector Vocabularies (PSV)} cover public sector concepts, the \emph{Medical Subject Headings (MeSH)} and \emph{Peroxisome Knowledge Base (PXV)} are from the (bio-)medical domain, the \emph{Geonames Ontology} covers locations, the \emph{Thesaurus for Economics (STW)} and \emph{North American Industry Classification System (NAICS)} are vocabularies from the economics and business area and the \emph{DBPedia Categories} reflect Wikipedia's user-generated categorization system. We also include the \emph{Meketre} vocabulary, which defines Egyptology-related concepts, and the non-public \emph{Austrian Armed Forces Thesaurus (LVAk)}, which we generated from CSV files provided by the vocabulary maintainers. A zip-file containing all 14 public vocabulariesx we used in our analysis, can be downloaded from: \url{http://www.github.com/cmader/???}. In contrast to the column ``Concepts'' that shows the total number of SKOS Concepts that can be infered from the vocabulary, the column ``Authoritative Concepts`` lists only those concepts that have an URI in the namespace of the vocabulary.

@@

\textbf{Vocabulary} & \rotatebox{90}{\textbf{Abbreviation}} & \rotatebox{90}{\textbf{Version/last mod.}} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{Auth. Concepts}} & \rotatebox{90}{\textbf{Labels}} & \rotatebox{90}{\textbf{Semantic Rel.}} & \rotatebox{90}{\textbf{Aggregation Rel.}} & \rotatebox{90}{\textbf{Concept Schemes}} & \rotatebox{90}{\textbf{Collections}}\\
\toprule
Agricultural Thesaurus & \textbf{AGROVOC} & 1.3 & 32,035 & 32,035 & 620,629 & 65,934 & 32,085 & 1 & 0 \\
\hline
DBpedia Categories & \textbf{DBpedia} & ? & 743,410 & 743,410 & 740,352 & 1,490,316 & 0 & 0 & 0 \\
\hline
The EU's multilingual thesaurus & \textbf{Eurovoc} & 5.0 & 6,797 & 6,797 & 457,788 & 18,491 & 15,512 & 128 & 0 \\
\hline
Geonames Ontology & \textbf{Geonames} & 2.2.1 & 671 & 671 & 671 & 0 & 671 & 9 & 0 \\
\hline
Gemeenschappelijke Thesaurus Audiovisuele Archieven & \textbf{GTAA} & 2010/08/25 & 171,991 & 171,991 & 178,776 & 50,892 & 343,980 & 9 & 0 \\
\hline
Integrated Public Sector Vocabulary & \textbf{IPSV} & 2.00 & 4,732 && 7,945 & 13,843 & 4,483 & 3 & 0 \\
\hline
Library of Congress Subject Headings & \textbf{LCSH} & 2011/08/09 & 459,182 & 407,908 & 746,076 & 595,754 & 815,816 & 19 & 0 \\
\hline
Austrian Armed Forces Thesaurus & \textbf{LVAk} & n/a & 13,411 & 13,411 & 17,250 & 16,346 & 0 & 0 & 0 \\
\hline
Middle Kingdom tombs of Ancient Egypt Thesaurus & \textbf{Meketre} & 2011/07/07 & 422 & 422 & 569 & 1,698 & 6 & 2 & 0 \\
\hline
Medical Subject Headings & \textbf{MeSH} & ? & 24,626 & 24,626 & 150,617 & 38,858 & 0 & 0 & 0 \\
\hline
North American Industry Classification System & \textbf{NAICS} & 2012 & 4,175 & 2,213 & 0 & 8,684 & 2,235 & 1 & 0 \\
\hline
New York Times People & \textbf{NYTP} & 2010/06/22 & 4,979 & 4,979 & 4,979 & 0 & 4,979 & 1 & 0 \\
\hline
University of Southampton Pressinfo & \textbf{Pressinfo} & 2011/02/24 & 1,125 & 1,125 & 0 & 0 & 0 & 0 & 0 \\
\hline
Peroxisome Knowledge Base & \textbf{PXV} & 1.6 & 2,112 & 1,686 & 3,628 & 2,695 & 1,716 & 1 & 0 \\
\hline
Thesaurus for Economics & \textbf{STW} & 8.06 & 6,524 & 6,524 & 31,189 & 57,907 & 6,531 & 1 & 0 \\
\bottomrule
\end{tabular}

@@

Valueless Associative Relations & 9,438 & 0 & 495 & 0 & 1 && 282 & 8,120 & 0 & 0 & 1,879 & 0 & 5,082 & 0 & 5 \\

Solely Transitively Related Concepts & 0 & 0 & 0 & 0 & 2,652 && 0 & 0 & 0 & 0 & 0 & 36 & 0 & 2,189 & 0 \\

Omitted Top Concepts & 9 & 9 & 0 & 0 & 1 && 0 & 0 & 0 & 1 & 18 & 0 & 0 & 0 & 0 \\

Top Concept Having Broader Concepts & 0 & 0 & 0 & 1 & 0 && 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\

@@

Missing In-Links && 19 && 1,686 &&&&& 1,125 & 20 && 422 & 6,516 & 2,213 & 13,411 \\

Missing Out-Links & 171,991 & 671 && 1,472 & 6,797 && 32,035 & 743,410 & 1,116 & 0 & 408,198 & 273 & 6,524 & 0 & 13,411 \\

@@

% Labeling and Documentation Issues
We found labeling and documentation issues in all analyzed vocabularies. PXV provides labels for XX\% of its concepts, Pressinfo labels only for 1 out of 1125 concepts, and the STW thesaurus misses labels for two concepts. Eurovoc and NAICS use SKOS-XL labels, which are currently not supported by qSKOS.
Incomplete Language Coverage were issues in the three analyzed vocabularies: in AGROVOC all concepts had incomplete coverage, probably because \todo{CM}{es gibt labels in insgesamt 25 versch sprachen und keines der konzepte ist in jeder sprache beschrieben}, and in STW we found XX\%.
Most vocabularies didn't document their concepts with the documentary properties defined by SKOS.
Label conflicts, hence \todo{CM}{explain briefly} were found in several vocabularies: GTAA has many because? PXV has 7. Why given examples?

% Structural Issues

% Linked Data Specific Issues

>>>>

@@

We used Sindice and available listings\footnote{\url{http://www.w3.org/2001/sw/wiki/SKOS/Datasets}} to learn about existing SKOS vocabularies. Table~\ref{tab:vocabs} lists our representative vocabulary selection and shows that the vocabularies differ in size and application domain: the \emph{Gemeenschappelijke Thesaurus Audiovisuele Archieven (GTAA)} is vocabulary from the media domain, \emph{Eurovoc} and the \emph{Integrated Public Sector Vocabularies (PSV)} cover public sector concepts, the \emph{Medical Subject Headings (MeSH)} and \emph{Peroxisome Knowledge Base (PXV)} are from the (bio-)medical domain, the \emph{Geonames Ontology} covers locations, the \emph{Thesaurus for Economics (STW)} and \emph{North American Industry Classification System (NAICS)} are vocabularies from the economics and business area and the \emph{DBPedia Categories} reflect Wikipedia's user-generated categorization system. We also include the \emph{Meketre} vocabulary, which defines Egyptology-related concepts, and the non-public \emph{Austrian Armed Forces Thesaurus (LVAk)}, which we generated from CSV files provided by the vocabulary maintainers. A zip-file containing all 14 public vocabulariesx we used in our analysis, can be downloaded from: \url{http://www.github.com/cmader/???}.

@@

\textbf{Vocabulary} & \rotatebox{90}{\textbf{Abbreviation}} & \rotatebox{90}{\textbf{Version/last mod.}} & \rotatebox{90}{\textbf{Concepts}} & \rotatebox{90}{\textbf{Auth. Concepts}} & \rotatebox{90}{\textbf{Labels}} & \rotatebox{90}{\textbf{Semantic Rel.}} & \rotatebox{90}{\textbf{Aggregation Rel.}} & \rotatebox{90}{\textbf{Concept Schemes}} \\
\toprule
Agricultural Thesaurus & \textbf{AGROVOC} & 1.3 & 32,035 & 32,035 & 620,629 & 65,934 & 32,085 & 1 \\
\hline
DBpedia Categories & \textbf{DBpedia} & ? & 743,410 & 743,410 & 740,352 & 1,490,316 & 0 & 0 \\
\hline
The EU's multilingual thesaurus & \textbf{Eurovoc} & 5.0 & 6,797 & 6,797 & 457,788 & 18,491 & 15,512 & 128 \\
\hline
Geonames Ontology & \textbf{Geonames} & 2.2.1 & 671 & 671 & 671 & 0 & 671 & 9 \\
\hline
Gemeenschappelijke Thesaurus Audiovisuele Archieven & \textbf{GTAA} & 2010/08/25 & 171,991 & 171,991 & 178,776 & 50,892 & 343,980 & 9 \\
\hline
Integrated Public Sector Vocabulary & \textbf{IPSV} & 2.00 & 4,732 & 3,080 & 7,945 & 13,843 & 4,483 & 3 \\
\hline
Library of Congress Subject Headings & \textbf{LCSH} & 2011/08/09 & 459,182 & 407,908 & 746,076 & 595,754 & 815,816 & 19 \\
\hline
Austrian Armed Forces Thesaurus & \textbf{LVAk} & n/a & 13,411 & 13,411 & 17,250 & 16,346 & 0 & 0 \\
\hline
Middle Kingdom tombs of Ancient Egypt Thesaurus & \textbf{Meketre} & 2011/07/07 & 422 & 422 & 569 & 1,698 & 6 & 2 \\
\hline
Medical Subject Headings & \textbf{MeSH} & ? & 24,626 & 24,626 & 150,617 & 38,858 & 0 & 0 \\
\hline
North American Industry Classification System & \textbf{NAICS} & 2012 & 4,175 & 2,213 & 0 & 8,684 & 2,235 & 1 \\
\hline
New York Times People & \textbf{NYTP} & 2010/06/22 & 4,979 & 4,979 & 4,979 & 0 & 4,979 & 1 \\
\hline
University of Southampton Pressinfo & \textbf{Pressinfo} & 2011/02/24 & 1,125 & 1,125 & 0 & 0 & 0 & 0 \\
\hline
Peroxisome Knowledge Base & \textbf{PXV} & 1.6 & 2,112 & 1,686 & 3,628 & 2,695 & 1,716 & 1 \\
\hline
Thesaurus for Economics & \textbf{STW} & 8.06 & 6,524 & 6,524 & 31,189 & 57,907 & 6,531 & 1 \\
\bottomrule
\end{tabular}

@@

Valueless Associative Relations & 9,438 & 0 & 495 & 0 & 1 & 239 & 282 & 8,120 & 0 & 0 & 1,879 & 0 & 5,082 & 0 & 5 \\

Solely Transitively Related Concepts & 0 & 0 & 0 & 0 & 2,652 & 0 & 0 & 0 & 0 & 0 & 0 & 36 & 0 & 2,189 & 0 \\

Omitted Top Concepts & 9 & 9 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 1 & 18 & 0 & 0 & 0 & 0 \\

Top Concept Having Broader Concepts & 0 & 0 & 0 & 1 & 0 &0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\

@@

Missing In-Links & 171,980* & 19 && 1,686 & 6,796 & 3,080 & 32,035 & 733,800* & 1,125 & 20 & 404,540* & 422 & 6,516 & 2,213 & 13,411 \\

Missing Out-Links & 171,991 & 671 && 1,472 & 6,797 & 0 & 32,035 & 743,410 & 1,116 & 0 & 408,198 & 273 & 6,524 & 0 & 13,411 \\

@@

% Labeling and Documentation Issues
\todo{BH}{this sentence is not clear, I assume you mean language tags, not labels.I'm not sure if relating language tags to concepts makes sense because a concept can be connected to multiple labels}
We found labeling and documentation issues in all analyzed vocabularies. PXV provides labels for XX\% of its concepts, Pressinfo labels only for 1 out of 1125 concepts, and the STW thesaurus misses labels for two concepts. Eurovoc and NAICS use SKOS-XL labels, which are currently not supported by qSKOS.

Incomplete Language Coverage were issues in the three analyzed vocabularies: in AGROVOC all concepts had incomplete coverage, probably because \todo{CM}{es gibt labels in insgesamt 25 versch sprachen und keines der konzepte ist in jeder sprache beschrieben}, and in STW we found XX\%.

Most vocabularies didn't document their concepts with the documentary properties defined by SKOS.

Label conflicts, hence \todo{CM}{explain briefly} were found in several vocabularies: GTAA has many because? PXV has 7. Why given examples?

% Structural Issues
orphan concepts: seem to depend on vocab use case: list of names: many orphans (gtaa,lcsh,geonames); vocabs establishing a network of terms: less to none orphan concepts

wcc: "islands" of semantically related concepts; most interesting cases: vocabs having few orphan concepts but more than one wcc: vocab may split into clusters inintentionally

cycles: not a common issue, dbpedia category cylces were somewhat surprising

valueless associative relations: in 9 of 15 vocabs; comparatively few in relation to all semantic relations

solely transitively related concepts: only 3 of 15 vocabs

omitted top concepts: 11 vocabs have at least one concept scheme defined, 5 have missing top concepts; gtaa,geonames,nyt people have no top concepts for any concept scheme, eurovoc uses 128 concept schemes and only one without top concept; opposite case for lcsh: uses 19 concept schemes defined, only one has a top concept

top concepts having broader concept: only pxv vocab affected in current version: in earlier version a different concept featuring this issue could be spotted: according to vocab creator this was an erroneous concept still available in the triplestore; maybe introduced by some bug in the vocab management software

% Linked Data Specific Issues
missing in-links: authoritative conceps of most vocabs aren't externally linked; exception: nyt people vocab

missing out-links: many vocab's also don't provide links to other vocabs; exception: naics, nyt people

broken loinks: meketre: reason server-side misconfiguration (fixed in current version); dbpedia sometimes down, link checking for some vocabs difficult: our algorithm waits 3secs when dereferencing links to not produce an uninteded DOS attack => estimation was necessary

undef skos resources: seems to be a minor problems, the examined voabs seem to be well-maintained

### 2012-03-26 criteria.tex
added some keywords for results interpretation

% Origin and rationale
SKOS concepts become linkable resources, when vocabularies are being published on the Web. Estimating the number of in-links and identifying the concepts without any in-links, can be an important indicator for the importance of a concept. The required in-links for such an estimation can be obtained from existing Linked Data indices like, e.g., Sindice\footnote{\url{http://sindice.com/}}. 
% Function definition
The function \textit{mil} estimates the number of in-links referencing a concept, by iterating over all elements in $C$ and querying the Sindice SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicator for missing  in-links and are thus included in the result set of this function.

>>>>

% Origin and rationale
SKOS concepts become linkable resources, when vocabularies are being published on the Web. Estimating the number of in-links and identifying the concepts without any in-links, can be an important indicator for the importance of a concept. The required in-links for such an estimation can be obtained from existing Linked Data indices like, e.g., Sindice\footnote{\url{http://sindice.com/}}. 
% Function definition
To estimate the number of in-links we only take a set $AC \subseteq C$ into account which consists of all ``authoritative concepts'', i.e. concepts that have an URI in the namespace of the vocabulary. The function \textit{mil} estimates the number of in-links referencing an authoritative concept, by iterating over all elements in $AC$ and querying the Sindice SPARQL endpoint for triples containing the concept's URI in the object part. Empty query results are indicator for missing in-links and are thus included in the result set of this function.
